Extended Data

### 2021-10-06 doc/manuscript/Dow_manuscript_Nature.Rmd
edited tree-ring section
@camerondow35 , I dramatically shortened this, removing the separate detailed discussion of patterns at SCBI and HF.

I also started a short paragraph on the second analysis, which will need some details filled in.

To understand how annual growth increments have responded to spring temperatures at the centennial scale, we analyzed tree-ring chronologies from our focal study sites  (`r EDT_species`) and across an additional `r n_cores_sites-2` locations (Fig. `r tree_rings`, `r EDF_map`).
For `r n_chronologies_SCBI` species at SCBI and `r n_chronologies_HF` species at Harvard Forest, ring width indices (RWI) rarely exhibited significant correlations to either $T_{max}$ or $T_{min}$ in the months corresponding to the critical temperature windows identified here (Fig. `r tree_rings`, `r EDF_tree_rings_TMIN`).
Although correlations were rarely significant, there was a consistent pattern in responses to $T_{max}$ across all species at both sites:
responses to $T_{max}$ were negative or neutral in all months from February to August, with the exception of one spring month at each site (April at SCBI, May at Harvard Forest) where growth displayed a positive or neutral response to $T_{max}$ (Fig. `r tree_rings`).
In these months, positive correlations to $T_{max}$ were statistically significant (at p=0.05) for one of nine ring-porous species at SCBI and one of two diffuse-porous species at each site.
Starting in May at SCBI and June at Harvard Forest, negative correlations of RWI to $T_{max}$ were predominant, being statistically significant in at least one growing season month for all `r n_chronologies_SCBI` species at SCBI and one of `r n_chronologies_HF` species at Harvard Forest. 
Patterns in $T_{min}$ were fairly similar to those of $T_{max}$, but weaker and more stochastic, with more (non-significant) correlations going in opposite directions (`r EDF_tree_rings_TMIN`).

@@

The patterns observed at our focal sites were mirrored across an additional `r n_chronologies-n_chronologies_SCBI-n_chronologies_HF` chronologies from `r n_cores_sites-2` sites (Fig. `r tree_rings`; `r EDF_map`)[@dorangeville_drought_2018].
In total, our analysis included `r n_chronologies` species at sites distributed from Alabama (Lat = 34.35) to Michigan (Lat = 45.56) and spanning a 15 $^\circ$C range in April $T_{max}$.
Across all chronologies (focal sites included), RWI was significantly positively correlated with April $T_{max}$ for only 1% of records: 1 of 142 ring-porous and 2 of 66 diffuse-porous species-site combinations.
In contrast, RWI was frequently significantly negatively correlated with $T_{max}$ during peak growing season months (May-August), with significant correlations for 53% (May: 47/142, Jun: 108/142, Jul: 90/142, Aug: 57/142) and 46% (May: 11/66, Jun: 53/66, Jul: 37/66, Aug: 22/66) of species-site-month combinations for ring-porous and diffuse-porous species, respectively.
Again, $T_{min}$ generally exerted less influence over annual growth than $T_{max}$ (`r EDF_tree_rings_TMIN`), with few significant correlations between spring $T_{min}$ and RWI.

>>>>

To understand how annual growth increments have responded to spring temperatures at the centennial scale, we analyzed tree-ring chronologies of `r n_chronologies_SCBI` species at SCBI and `r n_chronologies_HF` species at Harvard Forest (`r EDT_species`), and an additional `r n_chronologies-n_chronologies_SCBI-n_chronologies_HF` chronologies from `r n_cores_sites-2` sites (Fig. `r tree_rings`; `r EDF_map`)[@dorangeville_drought_2018].
In total, our analysis included `r n_chronologies` species at sites distributed from Alabama (Lat = 34.35) to Michigan (Lat = 45.56) and spanning a 15 $^\circ$C range in April $T_{max}$.
Across all chronologies, RWI was significantly (at p=0.05) positively correlated with April $T_{max}$ for only 1% of records: 1 of 142 ring-porous and 2 of 66 diffuse-porous species-site combinations.
In contrast, RWI was frequently significantly negatively correlated with $T_{max}$ during peak growing season months (May-August), with significant correlations for 53% (May: 47/142, Jun: 108/142, Jul: 90/142, Aug: 57/142) and 46% (May: 11/66, Jun: 53/66, Jul: 37/66, Aug: 22/66) of species-site-month combinations for ring-porous and diffuse-porous species, respectively.
$T_{min}$ generally exerted less influence over annual growth than $T_{max}$, with few significant correlations between spring $T_{min}$ and RWI (`r EDF_tree_rings_TMIN`).

To test whether a positive effect of April temperatures might be offset by a negative effect of summer temperatures, we tested for their joint and interactive effects.
Results were qualitatively similar to the univariate correlations (Fig. `r tree_rings`), with a positive effect of April $T_{max}$ for only **#**% of records and significant negative correlations with summer $T_{max}$ for **##-##**% of records, depending on the set of summer months included in the analysis (**(*Appendix Table*)**).
Significant interactive effects between spring and summer temperatures were rare (**DETAILS**).

@@
